---
title: "Final Project (MBDA750)    "
author: "**Authors (Team Bravo): ** Jehu Humphries"
date: "2024"
output:
  html_notebook:
    pandoc_args: ["--from", "markdown+autolink_bare_uris"]
---
# SHAPING

## -----FrameWork

```{r, message=FALSE, warning=FALSE}
# Importing all necessary R libraries
library(caret)       # Classification and Regression Training
library(conflicted)  # Provides an alternative approach to reolving conflicts caused by ambiguous function names
library(corrplot)    # provides a visual exploratory tool on correlation matrix that supports automatic variable reordering 
library(dplyr)       # Used for data manipulation
library(e1071)       # Misc Functions of the DEPT of Statistics, Probability Theory Group
library(GGally)      # Adds functions to ggplot2 that simplify combining geometric objects with transformed data. 
library(ggcorrplot)  # Visualization of a correlation matrix using ggplot2
library(glm2)        # Fitting Generalized Linear Models
library(lubridate)   # Tools that make it easier to parse and manipulate dates
library(pROC)        # An R package to display and analyze ROC curves
library(reticulate)  # R Interface to Python
library(ROCR)        # Visualizing the Performance of Scoring Classifiers
library(rquery)      # Relational Query Generator for Data Manipulation at Scale
library(tidyverse)   # Contains dplyr, forcats, ggplot2, readr, tibble, tidyr
library(visdat)      # Preliminary Visualisation of Data
```

```{python}
# Python: Import the Packages above for Hybrid Analysis and Plots
import matplotlib.pyplot as plt
import numpy as np
import pandas as pd
import plotly.express as px
import seaborn as sns
from sklearn.metrics import classification_report
from sklearn.metrics import confusion_matrix
from sklearn.metrics import accuracy_score
import warnings 
warnings.filterwarnings('ignore')
```

```{r}
# Create 
CovDths <- read.csv("CovidDeaths.csv")
CovDths$date <- as.Date(CovDths$date, format="%m/%d/%Y")
View(CovDths)
```

```{r}
# Create 
CovVac <- read.csv("CovidVaccinations.csv")
CovVac$date <- as.Date(CovVac$date, format="%m/%d/%Y")
View(CovVac)
```

## -----Exploration

### -----CovidDeaths

```{r}
dim(CovDths)
```

```{r}
sum(is.na(CovDths))
```

```{r}
# Show Variable titles and data types. 
str(CovDths)
```

```{r}
# Show dataframe 'CovDths' variable characteristics. 
summary(CovDths)
```

```{r}
vis_dat(CovDths, warn_large_data = FALSE)
```

### -----CovidVaccinations

```{r}
dim(CovVac)
```

```{r}
sum(is.na(CovVac))
```

```{r}
# Show Variable titles and data types. 
str(CovVac)
```

```{r}
# Show variable characteristics. 
summary(CovVac)
```

```{r}
vis_dat(CovVac, warn_large_data = FALSE)
```

## -----Shaping (PCovDth)

```{python}
# Python: Created Pandas PAAd1 dataframe from PAdultC.csv for Association Analysis
PCovDth = pd.read_csv('CovidDeaths.csv')
```

```{python}
# Python: Gathered data on titles and column positions to prepare for dataframe shaping
column_positions = [PCovDth.columns.get_loc(col) for col in PCovDth.columns]
column_titles = PCovDth.columns.tolist()

for position, title in zip(column_positions, column_titles):
    print(f"Column Position: {position}, Title: {title}")
```

```{python}
PCovDth1 = PCovDth.iloc[:, [0, 1, 2, 3, 4, 5, 6, 8, 9, 11, 12, 14, 15, 17]]
PCovDth1.to_csv('PCovDth1.csv', index=False)
```

```{r}
CovDth1 <- read.csv("PCovDth1.csv")
vis_dat(CovDth1, warn_large_data = FALSE)
```

## -----Shaping (PCovVac)

```{python}
# Python: Created Pandas PAAd1 dataframe from PAdultC.csv for Association Analysis
PCovVac = pd.read_csv('CovidVaccinations.csv')
```

```{python}
# Python: Gathered data on titles and column positions to prepare for dataframe shaping
column_positions = [PCovVac.columns.get_loc(col) for col in PCovVac.columns]
column_titles = PCovVac.columns.tolist()

for position, title in zip(column_positions, column_titles):
    print(f"Column Position: {position}, Title: {title}")
```

```{python}
PCovVac1 = PCovVac.iloc[:, [4, 5, 6, 7, 10, 11, 13, 14, 15, 17, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40]]
PCovVac1.to_csv('PCovVac1.csv', index=False)
```

```{r}
CovVac1 <- read.csv("PCovVac1.csv")
vis_dat(CovVac1, warn_large_data = FALSE)
```

## -----Combining (CovDthVac)

```{r}
# Combine
CovDthVac <- cbind(CovDth1, CovVac1)
write.csv(CovDthVac, "CovDthVac.csv", row.names = FALSE)
colnames(CovDthVac)
```

```{r}
dim(CovDthVac)
```


```{r}
sum(is.na(CovDthVac))
```

```{r}
CovDthVac[is.na(CovDthVac)] <- 0                                                # Replace 'N/A' values in dataset with '0'
```

```{r}
sum(is.na(CovDthVac))
```

```{r}
png("CorPlot5.png", width = 1200, height = 800)
vis_dat(CovDthVac, warn_large_data = FALSE)
dev.off()
```

```{r}
summary(CovDthVac)
```

## -----Linear Relationships (Correlation Matrix)

### -----Rquery.cormat()

```{r}
NumCols <- CovDthVac %>% select_if(is.numeric)                                  # Select only the numeric columns
CovDthVacNum <- data.frame(NumCols)                                             # Create a new dataframe with only the numeric columns
summary(CovDthVacNum)                                                               # Show DataFrame data types
```

```{r}
CorMtrx <- cor(CovDthVacNum)                                                    # Compute the correlation matrix
png("CorPlot1.png", width = 1200, height = 800)
corrplot::corrplot(CorMtrx, method = "circle", type = "full", tl.col = "black") # Plot the correlation matrix
title("Rquery.Cormat Correlation Matrix of CovDthVacNum Data Frame")
dev.off()
```

```{r}
corrplot::corrplot(CorMtrx, method = "circle", type = "full", tl.col = "black") # Plot the correlation matrix
title("Rquery.Cormat Correlation Matrix of CovDthVacNum Data Frame")
```

```{r}
CorMtrx <- cor(CovDthVacNum)                                                    # Compute the correlation matrix
png("CorPlot2.png", width = 1200, height = 800)
ggcorrplot::ggcorrplot(CorMtrx, method = "circle", type = "full", tl.col = "black") + # Plot the correlation matrix
  ggplot2::labs(title = "Rquery.Cormat Matrix of CovDthVacNum Data Frame")
dev.off()
```

```{r}
ggcorrplot::ggcorrplot(CorMtrx, method = "circle", type = "full", tl.col = "black") + # Plot the correlation matrix
  ggplot2::labs(title = "Rquery.Cormat Matrix of CovDthVacNum Data Frame")
```

### -----Spearman Cor()

```{r}
CorMtrx2 <- cor(CovDthVacNum, method = "spearman")                              # Compute the Spearman correlation matrix
corrplot::corrplot(CorMtrx2, method = "circle", type = "full", tl.col = "black") # Plot the correlation matrix
title("Spearman Correlation Matrix of CovDthVacNum Data Frame")
```


```{r}
png("CorPlot3.png", width = 1200, height = 800)
corrplot::corrplot(CorMtrx2, method = "circle", type = "full", tl.col = "black")  # Plot the correlation matrix
title("Spearman Correlation Matrix of CovDthVacNum Data Frame")
dev.off()
```

```{r}
ggcorrplot::ggcorrplot(CorMtrx2, method = "circle", type = "full", tl.col = "black") + # Plot the correlation matrix
   ggplot2::labs(title = "Spearman Correlation Matrix of CovDthVacNum Data Frame")
```

```{r}
png("CorPlot4.png", width = 1200, height = 800)
ggcorrplot::ggcorrplot(CorMtrx2, method = "circle", type = "full", tl.col = "black") + # Plot the correlation matrix
  ggplot2::labs(title = "Spearman Correlation Matrix of CovDthVacNum Data Frame")
dev.off()
```

## -----Data HoldOut

```{r}
# Create DataFrame of 10% sample from CovDthVac called RdMt10. The remaining 90% is a DataFrame called RdMt90. Show Dimensions of each.
RdMt10 <- CovDthVac %>% sample_frac(0.1)
RdMt90 <- CovDthVac %>% anti_join(RdMt10)
dim(RdMt10)
dim(RdMt90)
```

## -----Data Partioning

```{r}
set.seed(123)                          # Set seed for reproducibility
RowIndic <- 1:nrow(RdMt90)             # Create a vector of row indices for RdMt90
TrainIndic <- createDataPartition(RowIndic, p = 0.7, list = FALSE)
RdMtTrain <- RdMt90[TrainIndic, ]      # Use createDataPartition() to split the indices into training and validation sets
RdMtValid <- RdMt90[-TrainIndic, ]
dim(RdMtTrain)
dim(RdMtValid)
```

## -----GLM2 Modeling 

```{r}
str(RdMtTrain)
```


```{r}
summary(RdMtTrain)
```

# -----Multivariate LogReg


```{r, warning=FALSE}
LogReg <- glm2(positive_rate ~ population + new_cases + new_deaths + people_fully_vaccinated + population_density +
                 median_age + gdp_per_capita + extreme_poverty + cardiovasc_death_rate + diabetes_prevalence + 
                 handwashing_facilities + hospital_beds_per_thousand + life_expectancy, family = binomial(link = "logit"), data = RdMtTrain)
summary(LogReg)
```

```{r}
# LR <- predict(LogReg, newdata=RdMtValid, type='response')
# p.readmitted <- round(LR)
# p.readmitted

# confusionMatrix(as.factor(p.readmitted), as.factor(RdMtValid$positive_rate))
```

```{r}
# Use model to predict probability of default
# predicted <- predict(model, test, type = "response")

# Fit logistic regression model
# model <- glm(positive_rate ~ ., family = "binomial", data = train)

# Use model to predict probability of default
# predicted <- predict(LogReg, RdMtValid, type = "response")

# Convert defaults from "Yes" and "No" to 1's and 0's
# test$positive_rate <- ifelse(test$positive_rate == "Yes", 1, 0)

# Create confusion matrix
# confusionMatrix(as.factor(RdMtValid$positive_rate), as.factor(predicted))
```

# -----LogReg2

```{r}
# Combine
CovDthVac2 <- cbind(CovDth1, CovVac1)
write.csv(CovDthVac, "CovDthVac2.csv", row.names = FALSE)
colnames(CovDthVac2)
```

```{r}
dim(CovDthVac2)
```

```{r}
# Remove 'N/A' rows in 'positive_rate' variable for Multivariate (Full-Fitted) Linear Regression
CovDthVac3 <- CovDthVac2 %>% dplyr::filter(!is.na(positive_rate))
```

```{r}
CovDthVac4 <- CovDthVac3 %>% mutate(positive_rate = ifelse(positive_rate > 0, 1, positive_rate))
```

```{r}
# Create new column PosBin
CovDthVac3 <- CovDthVac3 %>% mutate(PosBin = ceiling(positive_rate))
write.csv(CovDthVac3, "CovDthVac3.csv", row.names = FALSE)
```

```{r}
summary(CovDthVac3$PosBin)
```

```{r}
sum(CovDthVac3$PosBin == 0)
```

```{r}
dim(CovDthVac3)
```

```{r}
# Create new variable PosBin2
CovDthVac3 <- CovDthVac3 %>% mutate(PosBin2 = ifelse(positive_rate > 0.1, 1, 0))
```

```{r}
sum(CovDthVac3$PosBin2 == 0)
```

```{r, warning=FALSE}
LogReg1 <- glm2(positive_rate ~ population + new_cases + new_deaths + people_fully_vaccinated + population_density +
                 median_age + gdp_per_capita + extreme_poverty + cardiovasc_death_rate + diabetes_prevalence + 
                 handwashing_facilities + hospital_beds_per_thousand + life_expectancy, family = binomial(link = "logit"), data = CovDthVac3)
summary(LogReg1)
```

# YEP--LR for PosBin2

## --HoldOut2

```{r}
# Create DataFrame of 10% sample from CovDthVac called RdMt10. The remaining 90% is a DataFrame called RdMt90. Show Dimensions of each.
CovHold10 <- CovDthVac3 %>% sample_frac(0.1)
CovHold90 <- CovDthVac3 %>% anti_join(RdMt10)
dim(CovHold10)
dim(CovHold90)
```

## --Patitioning2

```{r}
set.seed(123)                          # Set seed for reproducibility
RowIndic <- 1:nrow(CovHold90)             # Create a vector of row indices for RdMt90
TrainIndic <- createDataPartition(RowIndic, p = 0.7, list = FALSE)
CovHoldTrain <- CovHold90[TrainIndic, ]      # Use createDataPartition() to split the indices into training and validation sets
CovHoldValid <- CovHold90[-TrainIndic, ]
dim(CovHoldTrain)
dim(CovHoldValid)
```
## ---LogRegTrain
```{r, warning=FALSE}
LogReg2 <- glm2(PosBin2 ~ population + new_cases + new_deaths + people_fully_vaccinated + population_density +
                 median_age + gdp_per_capita + extreme_poverty + cardiovasc_death_rate + diabetes_prevalence + 
                 handwashing_facilities + hospital_beds_per_thousand + life_expectancy, family = binomial(link = "logit"), data = CovHoldTrain)
summary(LogReg2)
```

```{r}
LR <- predict(LogReg2, newdata=CovHoldTrain, type='response')
p.readmitted <- round(LR)
p.readmitted

confusionMatrix(as.factor(p.readmitted), as.factor(CovHoldTrain$PosBin2))
```

```{r}
auc(CovHoldTrain$PosBin2, LR)
```

```{r}
roc <- roc(CovHoldTrain$PosBin2, LR)
plot(roc, main = "ROC Curve", print.auc = TRUE, auc.polygon = TRUE, max.auc.polygon = TRUE, grid = TRUE, legacy.axes = TRUE)
```



## ---LogRegValidate

```{r, warning=FALSE}
LogReg3 <- glm2(PosBin2 ~ population + new_cases + new_deaths + people_fully_vaccinated + population_density +
                 median_age + gdp_per_capita + extreme_poverty + cardiovasc_death_rate + diabetes_prevalence + 
                 handwashing_facilities + hospital_beds_per_thousand + life_expectancy, family = binomial(link = "logit"), data = CovHoldValid)
summary(LogReg3)
```

```{r}
LR1 <- predict(LogReg3, newdata=CovHoldValid, type='response')
p.readmitted <- round(LR1)
p.readmitted

confusionMatrix(as.factor(p.readmitted), as.factor(CovHoldValid$PosBin2))
```

```{r}
auc(CovHoldValid$PosBin2, LR1)
```

```{r}
roc <- roc(CovHoldValid$PosBin2, LR1)
plot(roc, main = "ROC Curve", print.auc = TRUE, auc.polygon = TRUE, max.auc.polygon = TRUE, grid = TRUE, legacy.axes = TRUE)
```

## ---LogRegScore

```{r, warning=FALSE}
LogReg4 <- glm2(PosBin2 ~ population + new_cases + new_deaths + people_fully_vaccinated + population_density +
                 median_age + gdp_per_capita + extreme_poverty + cardiovasc_death_rate + diabetes_prevalence + 
                 handwashing_facilities + hospital_beds_per_thousand + life_expectancy, family = binomial(link = "logit"), data = CovHold10)
summary(LogReg4)
```

```{r}
LR2 <- predict(LogReg4, newdata=CovHold10, type='response')
p.readmitted <- round(LR2)
p.readmitted

confusionMatrix(as.factor(p.readmitted), as.factor(CovHold10$PosBin2))
```

```{r}
auc(CovHold10$PosBin2, LR2)
```

```{r}
roc <- roc(CovHold10$PosBin2, LR2)
plot(roc, main = "ROC Curve", print.auc = TRUE, auc.polygon = TRUE, max.auc.polygon = TRUE, grid = TRUE, legacy.axes = TRUE)
```

# WK4 - AVS/PCA

### ---Section AVS

```{r, warning=FALSE, message=FALSE}
library(ade4)                 #         Tools for Multivariate Analysis
library(cluster)              # **NEW** "Finding Groups in Data": Cluster Analysis Extended Rousseeuw et al.
library(factoextra)           # **NEW** Print method for an object of class factoextra
library(FactoMineR)           # **NEW** Multivariate Exploratory Data Analysis and Data Mining with R
library(MASS)                 #         Support Functions and Datasets for Venables and Ripley's MASS 
library(NbClust)              # **NEW** Determining the Best Number of Clusters in a Data Set
library(PerformanceAnalytics) #         Econometric Tools for Performance and Risk Analysis
library(stats)                #         Functions for statistical calculations and random number generation
```

```{r}
# Create Data
CovHoldTrain0 <- CovHoldTrain
CovHoldTrain0[is.na(CovHoldTrain0)] <- 0
```

```{r}
dim(CovHoldTrain)
```

```{r}
sum(is.na(CovHoldTrain0))
```

```{r}
summary(CovHoldTrain0)
```


```{r, warning=FALSE}
LogReg5 <- glm2(PosBin2 ~ population + new_cases + new_deaths + people_fully_vaccinated + population_density +
                 median_age + gdp_per_capita + extreme_poverty + cardiovasc_death_rate + diabetes_prevalence + 
                 handwashing_facilities + hospital_beds_per_thousand + 
                 life_expectancy, family = binomial(link = "logit"), data = CovHoldTrain0)
summary(LogReg5)
```

```{r, warning=FALSE}
# Perform AVS using stepAIC
fit2 <- stepAIC(LogReg5, direction = "both")
summary(fit2)
```

```{r}
show(fit2)
```

```{r}
# Get the selected coefficients
selected_coeffs <- coef(fit2)
show(selected_coeffs)
```

```{r}
summary(selected_coeffs)
```


```{r}
# Now, if you want to predict using the selected variables, you can do the following:
# Assuming you have a new data frame 'newdata' with the same predictor variables
predicted_y <- predict(fit2, newdata = CovHoldTrain0)

# If you're interested in the names of the selected variables:
selected_variable_names <- attr(terms(fit2), 'term.labels')
show(selected_variable_names)
# AI-generated code. Review and use carefully. More info on FAQ.
# Remember to replace fit1 with your actual fitted model and adjust the variable names accordingly. The stepAIC() function will automatically select the best subset of variables based on AIC, and you can use the resulting model for predictions or further analysis123.
```

### ---Section PCA

```{r}
# DataFrame (Numerical Variables Only) CovDthVacNum summary (N/A Variables replaced with '0')
summary(CovDthVacNum) 
```

```{r}
# Create new dataframe only highest correlations from Multivariate (Full-Fit) Linear Regression (WK3)
CovDthVac5 <- CovDthVac3[, c("PosBin2" ,"population", "new_cases", "new_deaths", "median_age", "gdp_per_capita", "life_expectancy")]
```

```{r}
# Replace N/A values in DataFrame CovDthVac5 with '0'
CovDthVac5[is.na(CovDthVac5)] <- 0
```

```{r}
CorMtrx1 <- cor(CovDthVac5)                                                    # Compute the correlation matrix
png("CorPlot1.png", width = 1200, height = 800)
corrplot::corrplot(CorMtrx1, method = "circle", type = "full", tl.col = "black") # Plot the correlation matrix
title("Rquery.Cormat Correlation Matrix of CovDthVacNum Data Frame")
dev.off()
```

```{r} 
corrplot::corrplot(CorMtrx1, method = "circle", type = "full", tl.col = "black") # Plot the correlation matrix
title("Spearman Correlation Matrix of CovDthVacNum Data Frame")
```

```{r}
# Show new approved Variable names and their column Indecies
VarName0 <- colnames(CovDthVacNum)

for (i in seq_along(VarName0)) {
  cat("Var Index", i, ":", VarName0[i], "\n")
}

```

```{r}
# Perform PCA using the princomp(). 'Cor' set to FALSE due to 
PCA1 <- princomp(CovDthVacNum, cor=FALSE)
print(PCA1)
```

```{r, warning=FALSE}
# Producing a biplot from the princomp() results
biplot(PCA1)
```

```{r}
# Perform PCA using the prcomp()
PCA2 <- prcomp(CovDthVacNum, scale=TRUE)
PCA2$rotation <- -1*PCA2$rotation
PCA2$rotation
```

```{r}
# Perform PCA using the prcomp()
PCA2 <- prcomp(CovDthVacNum, scale=TRUE)
PCA2
```

```{r, warning=FALSE}
# Perform PCA using the dudi.pca()
PCA3 <- dudi.pca(CovDthVacNum)
PCA3
```

### -----FactoMineR

```{r}
# Strip PC1 variables below 0
ColDel <- c("total_cases_per_million", "new_cases_per_million", "total_deaths_per_million", "new_deaths_per_million",
            "reproduction_rate", "total_tests", "new_tests", "total_tests_per_thousand", "new_tests_per_thousand",
            "positive_rate", "tests_per_case", "stringency_index", "population_density", "median_age", "aged_65_older",
            "aged_70_older", "gdp_per_capita", "cardiovasc_death_rate")
CVDNStrip <- CovDthVacNum[, !(names(CovDthVacNum) %in% ColDel)]  # Create CVDNStrip by excluding the specified columns
dim(CVDNStrip)
```

```{r}
# Show new approved Variable names and their column Indecies
VarName <- colnames(CVDNStrip)
for (i in seq_along(VarName)) {
  cat("Var Index", i, ":", VarName[i], "\n")
}
```


```{r}
# Perform PCA using FactoMineR package and function PCA()
PCA4 <- PCA(CVDNStrip)
summary(PCA4)
```

```{r}
print(PCA4)
```

```{r}
EigVal <- PCA4$eig
head(EigVal[, 1:2])
```



```{r}
VarVar <- PCA4$var$cos2
show(VarVar)
```


```{r}
barplot(EigVal[, 2], names.arg=1:nrow(EigVal), 
       main = "Variances",
       xlab = "Principal Components",
       ylab = "Percentage of variances",
       col ="steelblue")
# Add connected line segments to the plot
lines(x = 1:nrow(EigVal), EigVal[, 2], 
      type="b", pch=19, col = "red")
```

```{r}
fviz_screeplot(PCA4, ncp=10)
```



```{r}
head(PCA4$var$coord)
```

```{r}
PCA4$var$contrib
```

```{r, warning=FALSE}
plot(PCA4, choix = "var")
```

```{r}
png("CorPlot1.png", width = 1200, height = 800)
fviz_pca_var(PCA4, col.var="contrib") +
  theme_minimal()
dev.off()
```

```{r}
fviz_pca_var(PCA4, col.var="contrib")+
  theme_minimal()
```

```{r}
fviz_pca_var(PCA4, col.var="cos2")+
  theme_minimal()
```

# WK5 KMeans Clustering

```{r, warning=FALSE}
# Strip PC1 variables below 0
ColDel1 <- c("total_cases_per_million", "new_cases_per_million", "total_deaths_per_million", "new_deaths_per_million",
            "reproduction_rate", "total_tests", "new_tests", "total_tests_per_thousand", "new_tests_per_thousand",
            "positive_rate", "tests_per_case", "stringency_index", "population_density", "median_age", "aged_65_older",
            "aged_70_older", "gdp_per_capita")
CDVNStrip1 <- CovDthVacNum[, !(names(CovDthVacNum) %in% ColDel1)]  # Create CVDNStrip by excluding the specified columns
write.csv(CDVNStrip1, "CDVNStrip1.csv", row.names = FALSE)
dim(CDVNStrip1)
```

```{r}
# Show new approved Variable names and their column Indecies
VarName1 <- colnames(CDVNStrip1)
for (i in seq_along(VarName1)) {
  cat("Var Index", i, ":", VarName1[i], "\n")
}
```

```{r}
summary(CDVNStrip1)
```

```{r}
# create a plot of the number of clusters vs. the total within sum of squares on dataframe CVDNSTRIP1
# fviz_nbclust(CDVNStrip1, kmeans, method = "wss")
```

```{r}
# Check if all columns are numeric
all_numeric <- all(sapply(CDVNStrip1, is.numeric))

if (all_numeric) {
  print("All columns in CDVNStrip1 are numeric.")
} else {
  print("Not all columns in CDVNStrip1 are numeric.")
}
```

```{r}
# Example: K-means clustering
k.max <- 10  # Maximum number of clusters to consider
wss <- sapply(2:k.max, function(k) {
  kmeans_result <- kmeans(CDVNStrip1, centers = k, nstart = 10)
  kmeans_result$tot.withinss
})
```

```{r}
# Visualize the optimal number of clusters
#fviz_nbclust(wss, method = "wss", k.max = k.max,
#             barfill = "steelblue", barcolor = "steelblue",
#             linecolor = "steelblue", print.summary = TRUE)
```

```{r}
# Compute the dissimilarity matrix (distance matrix) and determine optimal number of clusters 
DissM <- dist(CDVNStrip1)
nb <- NbClust(DissM, distance = "euclidean", min.nc = 2, max.nc = 10, method = "kmeans")
OptClust <- nb$Best.nc[1]
CDVNBclust <- kmeans(CDVNStrip1, centers = OptClust)

```



```{r}
# K-Mean on CDVNStrip1 (4 Clusters / 5 nstrarts)
set.seed(1)
KM1 <- kmeans(CDVNStrip1, centers = 4, nstart = 5)
KM1
```

```{r}
#plot results of final k-means model
fviz_cluster(KM1, data = CDVNStrip1)
```

```{r}
# Assuming CDVNStrip1 is your existing dataframe
# Select the desired columns
CDVNStrip2 <- CDVNStrip1 %>%
  dplyr::select(extreme_poverty, cardiovasc_death_rate, diabetes_prevalence,
         female_smokers, male_smokers, handwashing_facilities,
         hospital_beds_per_thousand, life_expectancy, human_development_index)
```

```{r}
#make this example reproducible
set.seed(1)

#perform k-means clustering with k = 4 clusters
km <- kmeans(CDVNStrip2, centers = 3, nstart = 5)

#view results
km
```

```{r}
#plot results of final k-means model
fviz_cluster(km, data = CDVNStrip2)
```

```{r}
# Copy the 'continent' column from CovDthVac to CDVNStrip1
CDVNStrip3 <- CDVNStrip1
CDVNStrip3$iso_code <- CovDthVac$iso_code
```

```{r}
CDVNStrip3$iso_code <- as.numeric(as.factor(CDVNStrip3$iso_code))
```


```{r}
#make this example reproducible
set.seed(1)

#perform k-means clustering with k = 4 clusters
km <- kmeans(CDVNStrip3, centers = 10, nstart = 25)

#view results
km
```

```{r}
#plot results of final k-means model
fviz_cluster(km, data = CDVNStrip3)
```
# --- YEARS

```{r}
CDVYears <- CovDthVac
CDVYears$Year <- substr(CDVYears$date, start = nchar(CDVYears$date) - 3, stop = nchar(CDVYears$date))
CDVYears$Year <- as.numeric(CDVYears$Year)
str(CDVYears)
```

```{r}
CDVYears <- select_if(CDVYears, is.numeric)
str(CDVYears)
```

```{r}
CDVYears <- CDVYears %>%
  dplyr::select(new_cases, new_deaths, new_vaccinations, cardiovasc_death_rate, 
                diabetes_prevalence, female_smokers, male_smokers, life_expectancy, 
                human_development_index, Year)
write.csv(CDVYears, "CDVYears.csv", row.names = FALSE)
str(CDVYears)
```


# --- 2020

```{r}
CDVYears <- read.csv("CDVYears.csv")
```


```{r}
CDVY20 <- CDVYears
CDVY20 <- CDVY20[CDVY20$Year == 2020, ]
CDVY20 <- subset(CDVY20, select = -Year)
str(CDVY20)
```

```{r}
fviz_nbclust(CDVY20, kmeans, method = "wss")
```

```{r}
#make this example reproducible
set.seed(1)

#perform k-means clustering with k = 4 clusters
km <- kmeans(CDVY20, centers = 4, nstart = 10)

#view results
km
```

```{r}
#plot results of final k-means model
fviz_cluster(km, data = CDVY20)
```

```

# --- 2021

```{r}
CDVY21 <- CDVYears
CDVY21 <- CDVY21[CDVY21$Year == 2021, ]
summary(CDVY21$Year)
```

# --- 2022

```{r}
CDVY22 <- CDVYears
CDVY22 <- CDVY22[CDVY22$Year == 2022, ]
summary(CDVY22$Year)
```

# --- 2023

```{r}
CDVY23 <- CDVYears
CDVY23 <- CDVY23[CDVY23$Year == 2023, ]
summary(CDVY23$Year)
```

# -----// END //----------------------------------------------------------------









