---
title:  "MBDA730 - Assignment #5 - Qualitative/Quantitative Bias Measurement - Jehu Humphries"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
  theme: cyborg
  source_code: embed
  social: ["twitter", "facebook", "menu"]
---

Row {data-height=250}
--------------------------------------------------------------------------------

### **GitHub / Library / Assignment:**
- GIT: JSH-ASGNMT#3-MBDA730 v1 @ HumphriesJehu/JSH-UC-BA (Public)
- Library: Flex Dashboard, Papaja, RSConnect, Shiny, TidyText, TidyVerse
```{r, echo=FALSE}
library(papaja)
library(rsconnect)
library(shiny)
library(tidytext)
library(tidyverse)
```

1. Pick an open source data from the sources page in this week. 
2. Read the instruction "Guide to address bias in datasets". 
3. Follow the points of qualitative AND quantitative assessment of bias detection for the dataset you picked in #1. 
4. Summarize your assessment based on #3 and post it on your github page.  Please include: dataset link, what bias assessment technique you used, and analysis script and result (R or RMD) 

Row {data-height=450}
--------------------------------------------------------------------------------

### **DataSet Name / Link / Characterization:**
- Data Set Name: DHDS_-_Prevalence_of_Disability_Status_and_Types_by_Demographic_Groups__2020
- Data Set Link: [DOWNLOAD](https://data.cdc.gov/api/views/qjg3-6acf/rows.csv?accessType=DOWNLOAD)

- **Original: 7,168 OBS under 30 Variables**
```{r, cache=TRUE}
DHDSDis <- read.csv("https://data.cdc.gov/api/views/qjg3-6acf/rows.csv?accessType=DOWNLOAD")
```
- Condense, Write, and Characterize New DataSet. **Remove Redundant Columns**
- **Condensed: 7,168 OBS under 11 Variables**
```{r}
DHDSDis1 <- DHDSDis[,c(2,7,9,10,11,13,14,15,16,19,24)]
write.csv(DHDSDis1, "DHDS-Prevalence_of_Disability.csv")
str(DHDSDis1)
```

Row {data-height=200}
--------------------------------------------------------------------------------

### **Dataset Background:**
- Dataset is a Center for Disease Control (CDC) Disability and Health Data System (DHDS)generated dataset.
- Disability and Health Data System (DHDS) is an online source of state-level data on adults with disabilities. Users can access information on six functional disability types: cognitive (serious difficulty concentrating, remembering or making decisions), hearing (serious difficulty hearing or deaf), mobility (serious difficulty walking or climbing stairs), vision (serious difficulty seeing), self-care (difficulty dressing or bathing) and independent living (difficulty doing errands alone).
- This dataset was created in 2018 and was last updated in May of 2022. The data in the dataset covers the year of 2022. This dataset is a four year old well established statistical product generated by a primary healthcare advocate in the United States, the CDC. Data is collected at the state level through the DHDS online system.

Row {data-height=150}
--------------------------------------------------------------------------------

### **QUALLTATIVE Assessment of Bias in the DataSet:**
- This dataset was created in 2018 and was last updated in May of 2022. The data covers the year 2020.
- This dataset is a four year old well established statistical product generated by a primary healthcare advocate in the United States, the CDC. 
- This dataset includes high and low confidence factors to compensate for input inconsistencies likely aggregated at the state level.

**//END**




