---
title:  "MBDA730 - Week 7 Analytical Homework - Jehu Humphries"
output: 
  flexdashboard::flex_dashboard:
  orientation: rows
  source_code: embed
  social: ["twitter", "facebook", "menu"]
---

Row {data-height=300}
--------------------------------------------------------------------------------
```{r, echo=FALSE}
library(papaja)
library(rmarkdown)
library(rsconnect)
library(shiny)
library(tidytext)
library(tidyverse)
```
### **GitHub / Library / Data Privacy Lab:**
- **GitHub:** JSH-ASGNMT#3-MBDA730 v1 @ HumphriesJehu/JSH-UC-BA (Public)
- **Library:** Flex Dashboard, Papaja, RMarkDown, RSConnect, Shiny, TidyText, TidyVerse- 
- **Data Privacy Lab:** Datafly: A System for Providing Anonymity in Medical Data
- The concept behind 'Datafly' is to look beyond the obvious personal identifiers in a data set which could be used to identify a person. Many personal identifiers such as name, address, social security numbers, etc... are automatically removed from some datasets to preserve the privacy of individuals. DataFly is designed to look beyond these types of information and generalize, substitute, insert, or remove other potentially linkable information. 
- The CDC dataset used in this assignment is an aggregation of data with the potentially linkable information already removed. There are no personally identifiable markers remaining in this dataset. 
- For this dataset, DataFly would be applied to the data as it was being aggregated to ensure no identification markers were maintained in the data.


Row {data-height=250}
--------------------------------------------------------------------------------

### DataSet Name / Link / Characterization:
- Data Set Name: DHDS_-_Prevalence_of_Disability_Status_and_Types_by_Demographic_Groups__2020
- Data Set Link: [DOWNLOAD](https://data.cdc.gov/api/views/qjg3-6acf/rows.csv?accessType=DOWNLOAD)

- **7,168 OBS under 30 Variables**
```{r, cache=TRUE, eval=FALSE}
DHDSDis <- read.csv("https://data.cdc.gov/api/views/qjg3-6acf/rows.csv?accessType=DOWNLOAD")
```
- Condense, Write, and Characterize New DataSet. **Remove Redundant Columns**
- **7,168 OBS under 11 Variables**
```{r, eval=FALSE}
DHDSDis1 <- DHDSDis[,c(2,7,9,10,11,13,14,15,16,19,24)]
write.csv(DHDSDis1, "DHDS-Prevalence_of_Disability.csv")
str(DHDSDis1)
```

Row {data-height=250}
--------------------------------------------------------------------------------

### Dataset Background:
- Dataset is a Center for Disease Control (CDC) Disability and Health Data System (DHDS)generated dataset.
- Disability and Health Data System (DHDS) is an online source of state-level data on adults with disabilities. Users can access information on six functional disability types: cognitive (serious difficulty concentrating, remembering or making decisions), hearing (serious difficulty hearing or deaf), mobility (serious difficulty walking or climbing stairs), vision (serious difficulty seeing), self-care (difficulty dressing or bathing) and independent living (difficulty doing errands alone).
- This dataset was created in 2018 and was last updated in May of 2022. The data in the dataset covers the year of 2022. This dataset is a four year old well established statistical product generated by a primary healthcare advocate in the United States, the CDC. Data is collected at the state level through the DHDS online system.

Row {data-height=200}
--------------------------------------------------------------------------------

### Qualitative Assessment of Bias:
- This dataset was created in 2018 and was last updated in May of 2022. The data in the dataset covers the year of 2022. This dataset is a four year old well established statistical product generated by a primary healthcare advocate in the United States, the CDC. Data is collected at the state level through the DHDS online system. The dataset includes high and low confidence factors to 

- //END




